## dplyr - 1

Throughout this course you will use dplyr to analyze a data set of airline flight data containing flights that departed from Houston. This data is stored in a package called hflights.

```{r,message=FALSE,warning=FALSE}
# Load the dplyr package
require(dplyr)

# Load the hflights package
require(hflights)

# Call both head() and summary() on hflights
head(hflights)

summary(hflights)

# Number of Observations and columns
dim(hflights)
```

### Convert data.frame to table

A tbl is just a special kind of data.frame. They make your data easier to look at, but also easier to work with. On top of this, it is straightforward to derive a tbl from a data.frame structure using tbl_df().

The tbl format changes how R displays your data, but it does not change the data's underlying data structure. You can do anything with the hflights tbl that you could do with the hflights data.frame.

```{r,message=FALSE,warning=FALSE}
# Convert the hflights data.frame into a hflights tbl
hflights <- tbl_df(hflights)

# Display the hflights tbl
hflights

# Create the object carriers, this is to prove tbl behaves like Dataframe.
carriers <- hflights$UniqueCarrier
```

### Changing labels of hflights, part 1

A bit of cleaning would be a good idea since the UniqueCarrier variable of hflights uses a confusing code system. 

To do this, let's work with a lookup table, that comes in the form of a named vector. When you subset the lookup table with a character string (like the character strings in UniqueCarrier), R will return the values of the lookup table that correspond to the names in the character string. 
```{r,message=FALSE,warning=FALSE}
lut <- c("AA" = "American", "AS" = "Alaska", "B6" = "JetBlue", "CO" = "Continental", 
         "DL" = "Delta", "OO" = "SkyWest", "UA" = "United", "US" = "US_Airways", 
         "WN" = "Southwest", "EV" = "Atlantic_Southeast", "F9" = "Frontier", 
         "FL" = "AirTran", "MQ" = "American_Eagle", "XE" = "ExpressJet", "YV" = "Mesa")

# Add the Carrier column to hflights
hflights$Carrier <- lut[hflights$UniqueCarrier]

hflights
```

### Changing labels of hflights, part 2

Let's try a similar thing, but this time to change the labels in the CancellationCode column. 

```{r,message=FALSE,warning=FALSE}
# The lookup table
unique(hflights$CancellationCode)

head(as.data.frame(subset(hflights, CancellationCode == "")))

hflights[hflights$CancellationCode == "",]$CancellationCode <- "E"


lut <- c("A" = "carrier", "B" = "weather", "C" = "FFA", "D" = "security", "E" = "not cancelled")

# Add the Code column
hflights$Code <- lut[hflights$CancellationCode]

# Glimpse at hflights
glimpse(hflights)

```

## The Select Verb

To answer the simple question whether flight delays tend to shrink or grow during a flight, we can safely discard a lot of the variables of each flight. To select only the ones that matter, we can use select().

select(df, var1, var2)

select(df, 1:4, -2) = Will select first four columns except the second one.

```{r,message=FALSE,warning=FALSE}
# Print out a tbl with the four columns of hflights related to delay
select(hflights,ActualElapsedTime:DepDelay)

# Print out the columns Origin up to Cancelled of hflights
select(hflights,Origin:Cancelled)

# Answer to last question: be concise!
# Find the most concise way to select: columns Year up to and including DayOfWeek, columns ArrDelay up to and including Diverted

names(hflights)
select(hflights,Year:DayOfWeek,ArrDelay:Diverted)

```

### Helper functions for variable selection

dplyr comes with a set of helper functions that can help you select groups of variables inside a select() call:

* starts_with("X"): every name that starts with "X",
* ends_with("X"): every name that ends with "X",
* contains("X"): every name that contains "X",
* matches("X"): every name that matches "X", where "X" can be a regular expression,
* num_range("x", 1:5): the variables named x01, x02, x03, x04 and x05,
* one_of(x): every name that appears in x, which should be a character vector.

```{r,message=FALSE,warning=FALSE}
# Print out a tbl containing just ArrDelay and DepDelay
select(hflights,contains("Delay"))

# Use a combination of helper functions and variable names to print out only the UniqueCarrier, FlightNum, TailNum, Cancelled, and CancellationCode columns of hflights
select(hflights,UniqueCarrier, ends_with("Num"),starts_with("Cancel"))

# Find the most concise way to return the following columns with select and its helper functions: DepTime, ArrTime, ActualElapsedTime, AirTime, ArrDelay, DepDelay
select(hflights,ends_with("Time"),ends_with("Delay"))

# Finish select call so that ex1d matches ex1r
names(hflights)
ex1r <- hflights[c("TaxiIn", "TaxiOut", "Distance")]
ex1d <- select(hflights, starts_with("Taxi"),Distance)

# Finish select call so that ex2d matches ex2r
ex2r <- hflights[c("Year", "Month", "DayOfWeek", "DepTime", "ArrTime")]
ex2d <- select(hflights,Year:ArrTime,-DayofMonth)

# Finish select call so that ex3d matches ex3r
ex3r <- hflights[c("TailNum", "TaxiIn", "TaxiOut")]
ex3d <- select(hflights,TailNum,starts_with("Taxi"))

```

## The Mutate Verb

mutate() creates new columns which are added to a copy of the dataset. 
mutate(df, z = x + y)

```{r,message=FALSE,warning=FALSE}
# Add the new variable ActualGroundTime to a copy of hflights and save the result as g1.
g1 <- mutate(hflights,ActualGroundTime = ActualElapsedTime - AirTime)

# Add the new variable GroundTime to g1. Save the result as g2.
g2 <- mutate(g1,GroundTime = TaxiIn + TaxiOut)
identical(g2$GroundTime,g2$ActualGroundTime)

# Add the new variable AverageSpeed to g2. Save the result as g3.
g3 <- mutate(g2,AverageSpeed = Distance / AirTime * 60)

# Print out g3
g3

```

### Add multiple variables using mutate

mutate(my_df, x = a + b, y = x + c) ## Reuse of the newly created column

```{r,message=FALSE,warning=FALSE}
# Add a second variable loss_ratio to the dataset: m1
m1 <- mutate(hflights, loss = ArrDelay - DepDelay, loss_ratio = loss/DepDelay)

# Add the three variables as described in the third instruction: m2
m2 <- mutate(hflights,TotalTaxi = TaxiIn + TaxiOut,ActualGroundTime = ActualElapsedTime - AirTime, Diff = TotalTaxi - ActualGroundTime)

```

## The Filter Verb

### Logical operators

R comes with a set of logical operators that you can use inside filter():

* x < y, TRUE if x is less than y
* x <= y, TRUE if x is less than or equal to y
* x == y, TRUE if x equals y
* x != y, TRUE if x does not equal y
* x >= y, TRUE if x is greater than or equal to y
* x > y, TRUE if x is greater than y
* x %in% c(a, b, c), TRUE if x is in the vector c(a, b, c)


```{r,message=FALSE,warning=FALSE}
# All flights that traveled 3000 miles or more
filter(hflights,Distance>=3000)

# All flights flown by one of JetBlue, Southwest, or Delta
filter(hflights,hflights$UniqueCarrier %in% c('B6','WN','DL'))

# All flights where taxiing took longer than flying
filter(hflights,(TaxiIn+TaxiOut) > AirTime)

```

### Combining tests using boolean operators

R also comes with a set of boolean operators that you can use to combine multiple logical tests into a single test. These include & (and), | (or), and ! (not). Instead of using the & operator, you can also pass several logical tests to filter(), separated by commas.

filter(df, a > 0 & b > 0)
filter(df, a > 0, b > 0)

```{r,message=FALSE,warning=FALSE}
# All flights that departed before 5am or arrived after 10pm
filter(hflights, DepTime < 500 | ArrTime > 2200)

# All flights that departed late but arrived ahead of schedule
filter(hflights, DepDelay > 0 & ArrDelay < 0)

# All flights that were cancelled after being delayed
filter(hflights, DepDelay > 0 & Cancelled == 1)

```

### Combination of Filter, Mutate and Select

```{r,message=FALSE,warning=FALSE}
# Select the flights that had JFK as their destination: c1
c1 <- filter(hflights,Dest == 'JFK')

# Combine the Year, Month and DayofMonth variables to create a Date column: c2
c2 <- mutate(c1,Date = paste(Year,Month,DayofMonth,sep = '-'))

# Print out a selection of columns of c2
select(c2,Date,DepTime,ArrTime,TailNum)

```

## Arranging your data using arrange verb

arrange() can be used to rearrange rows according to any type of data. 

If you pass arrange() a character variable, for example, R will rearrange the rows in alphabetical order according to values of the variable. 

If you pass a factor variable, R will rearrange the rows according to the order of the levels in your factor (running levels() on the variable reveals this order).

```{r,message=FALSE,warning=FALSE}
# Definition of dtc
dtc <- filter(hflights, Cancelled == 1, !is.na(DepDelay))

# Arrange dtc by departure delays
arrange(dtc,DepDelay)

# Arrange dtc so that cancellation reasons are grouped
arrange(dtc,CancellationCode)

# Arrange dtc according to carrier and departure delays
arrange(dtc,UniqueCarrier,DepDelay)

# Arrange according to carrier and decreasing departure delays
arrange(hflights,UniqueCarrier,desc(DepDelay))

# Arrange flights by total delay (normal order).
arrange(hflights,DepDelay + ArrDelay)

```


## Summarise verb

summarise(), the last of the 5 verbs, follows the same syntax as mutate(), but the resulting dataset consists of a single row instead of an entire new column in the case of mutate().

```{r,message=FALSE,warning=FALSE}
# Print out a summary with variables min_dist and max_dist
summarise(hflights,min_dist = min(Distance), max_dist = max(Distance))

# Print out a summary with variable max_div which are diverted

summarise(filter(hflights,Diverted == 1),max_div = max(Distance))

# Remove rows that have NA ArrDelay: temp1
temp1 <- filter(hflights,!is.na(ArrDelay))

# Generate summary about ArrDelay column of temp1
summarise(temp1,earliest = min(ArrDelay),average = mean(ArrDelay), latest = max(ArrDelay), sd = sd(ArrDelay))

# Keep rows that have no NA TaxiIn and no NA TaxiOut: temp2
temp2 <- filter(hflights,!is.na(TaxiIn) & !is.na(TaxiOut))

# Print the maximum taxiing difference of temp2 with summarise()
summarise(temp2,max_taxi_diff = max(abs(TaxiIn - TaxiOut)))

```
