## Text Mining

What is Text Mining?

The Process of distilling actionable insights from text

### Text Mining WorkFlow

1. Problem Definition and Specific goals
2. Identify text to be collected
3. Text Organization
4. Feature Extraction
5. Analysis
6. Reach an Insight, conclusion.

### Semantic Parsing vs Bag of Words

Semantic Parsing: We care about word type and order. This creates features lots of features. Single word can be part of a sentence or a Noun or a proper noun etc. Hence there are 3 features, hence feature rich. Therefore Semantic parsing follows the tree structure.

Bag of words: Does not care about word type or order. Words are considered as tokens.

Sometimes we can find out the author's intent and main ideas just by looking at the most common words. At its heart, bag of words text mining represents a way to count terms, or n-grams, across a collection of documents. 

```{r,warning=FALSE,message=FALSE}
text <- "Text mining usually involves the process of structuring the input text. The overarching goal is, essentially, to turn text into data for analysis, via application of natural language processing (NLP) and analytical methods."

# Load qdap
require(qdap)

# Find the 10 most frequent terms: term_count
term_count <- freq_terms(text,10)

# Plot term_count
plot(term_count)

```

### Building Corpus from vector and dataframe

#### 1. Vector

```{r,warning=FALSE,message=FALSE}
# Import text data
require(RCurl)

URLFile <- getURL("https://raw.githubusercontent.com/diegonogare/DataScience/master/Text%20Mining/coffee.csv")
tweets <- read.csv(text = URLFile,stringsAsFactors = FALSE)

# View the structure of tweets
str(tweets)

# Print out the number of rows in tweets
nrow(tweets)

# Isolate text from tweets: coffee_tweets
coffee_tweets <- tweets$text

head(coffee_tweets)

```

Now you've loaded your text data as a vector called coffee_tweets. Your next step is to convert this vector containing the text data to a corpus.

There are two kinds of the corpus data type, the permanent corpus, PCorpus, and the volatile corpus, VCorpus. In essence, the difference between the two has to do with how the collection of documents is stored in your computer. In this course, we will use the volatile corpus, which is held in your computer's RAM rather than saved to disk, just to be more memory efficient.

To make a volatile corpus, R needs to interpret each element in our vector of text, coffee_tweets, as a document. And the tm package provides what are called Source functions to do just that! In this exercise, we'll use a Source function called VectorSource() because our text data is contained in a vector. The output of this function is called a Source object.

```{r,warning=FALSE,message=FALSE}
# Load tm
require(tm)

# Make a vector source: coffee_source
coffee_source <- VectorSource(coffee_tweets)

```

Now that we've converted our vector to a Source object, we pass it to another tm function, VCorpus(), to create our volatile corpus.

The VCorpus object is a nested list, or list of lists. At each index of the VCorpus object, there is a PlainTextDocument object, which is essentially a list that contains the actual text data (content), as well as some corresponding metadata (meta).

For example, to examine the contents of the second tweet in coffee_corpus, you'd subset twice. Once to specify the second PlainTextDocument corresponding to the second tweet and again to extract the first (or content) element of that PlainTextDocument:

```{r,warning=FALSE,message=FALSE}
# Make a volatile corpus: coffee_corpus
coffee_corpus <- VCorpus(coffee_source)

# Print out coffee_corpus
coffee_corpus

# Print data on the 15th tweet in coffee_corpus
coffee_corpus[[15]]

# Print the content of the 15th tweet in coffee_corpus
coffee_corpus[[15]][1]

```


#### 2. Dataframe

For Dataframe  there is a Source function called DataframeSource(). The DataframeSource() function treats the entire row as a complete document, so be careful you don't pick up non-text data like customer IDs when sourcing a document this way.

```{r,warning=FALSE,message=FALSE}
# Example_text df

example_text <- data.frame(num = c(1,2,3), 
                           Author1 = c("Text mining is a great time.",
                                       "Text analysis provides insights",
                                       "qdap and tm are used in text mining"),
                           Author2 = c("R is a great language",
                                       "R has many uses",
                                       "DataCamp is cool!"))

example_text[,2:3]
# Create a DataframeSource on columns 2 and 3: df_source
df_source <- DataframeSource(example_text[,2:3])

# Convert df_source to a corpus: df_corpus
df_corpus <- VCorpus(df_source)

# Examine df_corpus
df_corpus

```

### Cleaning and Pre-processing text

#### Common cleaning functions from tm

In bag of words text mining, cleaning helps aggregate terms. For example, it may make sense that the words "miner", "mining" and "mine" should be considered one term. 

Common preprocessing functions include:

1. tolower(): Make all characters lowercase
2. removePunctuation(): Remove all punctuation marks
3. removeNumbers(): Remove numbers
4. stripWhitespace(): Remove excess whitespace

```{r,warning=FALSE,message=FALSE}
# Create the object: text
text <- "<b>She</b> woke up at       6 A.M. It\'s so early!  She was only 10% awake and began drinking coffee in front of her computer."

# All lowercase
tolower(text)

# Remove punctuation
removePunctuation(text)

# Remove numbers
removeNumbers(text)

# Remove whitespace
stripWhitespace(text)

```

#### Cleaning with qdap

The qdap package offers other text cleaning functions. Each is useful in its own way and is particularly powerful when combined with the others.

1. bracketX(): Remove all text within brackets (e.g. "It's (so) cool" becomes "It's cool")
2. replace_number(): Replace numbers with their word equivalents (e.g. "2" becomes "two")
3. replace_abbreviation(): Replace abbreviations with their full text equivalents (e.g. "Sr" becomes "Senior")
4. replace_contraction(): Convert contractions back to their base words (e.g. "shouldn't" becomes "should not")
5. replace_symbol() Replace common symbols with their word equivalents (e.g. "$" becomes "dollar")


```{r,warning=FALSE,message=FALSE}
# Remove text within brackets
bracketX(text)

# Replace numbers with words
replace_number(text)

# Replace abbreviations
replace_abbreviation(text)

# Replace contractions
replace_contraction(text)

# Replace symbols with words
replace_symbol(text)

```

#### All about stop words

Often there are words that are frequent but provide little information. So you may want to remove these so-called stop words. Some common English stop words include "I", "she'll", "the", etc. In the tm package, there are 174 stop words on this common list.

Using the c() function allows you to add new words (separated by commas) to the stop words list.

For eg: all_stops <- c("word1", "word2", stopwords("en"))

Once you have a list of stop words that makes sense, you will use the removeWords() function on your text. removeWords() takes two arguments: the text object to which it's being applied and the list of words to remove.

```{r,warning=FALSE,message=FALSE}
# List standard English stop words
stopwords("en")

# Print text without standard stop words
removeWords(text, stopwords("en"))

# Add "coffee" and "bean" to the list: new_stops
new_stops <- c("coffee","bean",stopwords("en"))

# Remove stop words from text
removeWords(text,new_stops)

```

#### Intro to word stemming and stem completion

The tm package provides the stemDocument() function to get to a word's root. This function either takes in a character vector and returns a character vector, or takes in a PlainTextDocument and returns a PlainTextDocument.

For example,

stemDocument(c("computational", "computers", "computation"))
returns "comput" "comput" "comput"

But because "comput" isn't a real word, we want to re-complete the words so that "computational", "computers", and "computation" all refer to the same word, say "computer", in our ongoing analysis.

We can easily do this with the stemCompletion() function, which takes in a character vector and an argument for the completion dictionary. The completion dictionary can be a character vector or a Corpus object

```{r,warning=FALSE,message=FALSE}
require(tm)
# Create complicate
complicate <- c("complicated","complication","complicatedly")

# Perform word stemming: stem_doc
stem_doc <- stemDocument(complicate)

# Create the completion dictionary: comp_dict
comp_dict <- "complicate"

# Perform stem completion: complete_text 
complete_text <- stemCompletion(stem_doc,comp_dict)

# Print complete_text
complete_text

```

#### Word stemming and stem completion on a sentence

Let's consider the following sentence as our document for this exercise:

"In a complicated haste, Tom rushed to fix a new complication, too complicatedly."

This sentence contains the same three forms of the word "complicate" that we saw in the previous exercise. The difference here is that even if you called stemDocument() on this sentence, it would return the sentence without stemming any words.

This happens because stemDocument() treats the whole sentence as one word. In other words, our document is a character vector of length 1, instead of length n, where n is the number of words in the document. 

To solve this problem, 
1. we first remove the punctation marks with the removePunctuation()
2. We then strsplit() this character vector of length 1 to length n
3. unlist(), then proceed to stem and re-complete.

```{r,warning=FALSE,message=FALSE}
## text_data

text_data <- "In a complicated haste, Tom rushed to fix a new complication, too complicatedly."

# Remove punctuation: rm_punc
rm_punc <- removePunctuation(text_data)

# Create character vector: n_char_vec
n_char_vec <- unlist(strsplit(rm_punc, split = ' '))

# Perform word stemming: stem_doc
stem_doc <- stemDocument(n_char_vec)

# Print stem_doc
stem_doc

# Create the completion dictionary: comp_dict
comp_dict <- "complicate"

# Re-complete stemmed document: complete_doc
complete_doc <- stemCompletion(stem_doc,comp_dict)

# Print complete_doc
complete_doc

```

#### Apply preprocessing steps to a corpus

The tm package provides a special function tm_map() to apply cleaning functions to a corpus. Mapping these functions to an entire corpus makes scaling the cleaning steps very easy.

To save time (and lines of code) it's a good idea to use a custom function like the one displayed in the editor, since you may be applying the same functions over multiple corpora.

clean_corpus <- function(corpus){  
  corpus <- tm_map(corpus, content_transformer(replace_abbreviation))  
  corpus <- tm_map(corpus, removePunctuation)  
  corpus <- tm_map(corpus, removeNumbers)  
  corpus <- tm_map(corpus, removeWords, c(stopwords("en"), "coffee"))  
  corpus <- tm_map(corpus, content_transformer(tolower))  
  return(corpus)  
} 

tm package functions do not need content_transformer(), but base R and qdap functions do.

Apply your customized function to the tweet_corp: clean_corp
clean_corp <- clean_corpus(coffee_corpus)

```{r,warning=FALSE,message=FALSE}


```

